# Fecal Indicator Bacteria (FIBs) and Hydrology  

We don't even really know the best way to even start looking at these relationships, so this portion will contain some exploratory visualizations.  

So to start, I'm just going to pull out a few sites where I would expect to see some influence - sites that aren't in the bays, but are maybe around the edges. I'll hand-pick and hard-code, to figure out what this all even looks like, then figure out how to scale up.  

We have monthly precipitation data and hydrologic load data.  




```{r}
library(mapview)
library(sf)
library(tidyverse)
library(tmap)
library(tbeptools)
library(khroma)
library(corrplot)
```

## Data Import  


```{r}
# spatial and precip files
load(here::here("sourced_data", "combined_precip.RData"))
load(here::here("sourced_data", "combined_spatial.RData"))
```


### FIB data import

```{r}
fib_dat <- readRDS(here::here("fib-data", "tbeptools_importwqp.rds"))

fib_sub <- fib_dat |> 
    filter(yr >= 2009,
           class == "Estuary",
           var == "ecocci")

# get those attached to bay segments  
fib_baysegs <- fib_sub |> 
    st_as_sf(coords = c("Longitude", "Latitude"),
             crs = "WGS84",
             remove = FALSE) |>    # WGS84 is default for handheld GPS units
    st_transform(crs = st_crs(swfwmdpixels)) |> 
    st_intersection(bay_segs) |> 
    mutate(bay_segment = case_match(BAY_SEG,
                                    "Boca Ciega Bay" ~ "Remainder Lower Tampa Bay",
                                    "Manatee River" ~ "Remainder Lower Tampa Bay",
                                    .default = BAY_SEG))

fib_hydro <- left_join(fib_baysegs, combnd,
                       by = c("yr" = "year",
                              "mo" = "month",
                              "bay_segment"))

```

### Attach daily precipitation  

# SOMETHING WENT WRONG IN THIS JOIN, WHAT????  

Only processed daily precip data back to 2018, so there are lots of NAs in the new data frame.  

```{r}
fib_daily_prcp <- fib_baysegs |> 
    mutate(date = as.Date(SampleTime),
           Season = case_when(mo %in% c(6:9) ~ "Wet",
                              .default = "Dry")) |> 
    left_join(prcp_daily_bayseg,
              by = c("date", "BAY_SEG")) |> 
    filter(!is.na(inches))
```


## Entero by monthly hydro load  

```{r}
ggplot(fib_hydro, aes(x = hydro_load,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. hydrologic load; log-log scale",
         x = "Hydrologic Load",
         y = "Enterococcus",
         col = "Month")
```

Interesting..... it's not necessarily the monthly hydrologic loads that matter (though it's clear that the higher hydrologic loads occur in certain months). Could come down to precipitation in some previous time periods.  

Could also be that because the majority of stations are in the bay proper, they're diluted and that's driving the lack-of-relationship. Really need to break this all out by "in-bay" vs. "have some land near the station".  

### Lagged hydro load  

```{r}
ggplot(fib_hydro, aes(x = lag1_hydro,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. lagged hydrologic load; log-log scale",
         x = "Hydrologic Load in previous month",
         y = "Enterococcus",
         col = "Month")
```

Nothing new.  

First a little more exploration of entero concentrations by time of year.  

## Entero by monthly precip  

```{r}
ggplot(fib_hydro, aes(x = inches,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.3) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. Monthly Precip",
         x = "Precipitation (in)",
         y = "Enterococcus",
         col = "Month")
```


We see it a bit more with precip rather than hydrologic load, at least in the higher values.  


### Previous month's precip  


```{r}
ggplot(fib_hydro, aes(x = lag1_prcp,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.3) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. Lagged Precip",
         x = "Precipitation (in) in previous month",
         y = "Enterococcus",
         col = "Month")
```

Not very different.  


## Entero by month  

```{r}
ggplot(fib_hydro,
       aes(x = factor(mo),
           y = val)) +
    geom_jitter(aes(col = factor(mo)),
                size = 0.7,
                alpha = 0.4) +
    geom_boxplot(alpha = 0) +
    scale_y_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero counts by month",
         x = "Month",
         y = "Enterococcus") +
    theme(legend.position = "none")
```

## Hydro load by month  

Only since 2009, to keep data frame size under control.  

```{r}
hydro_sub <- hydro |> 
    dplyr::filter(year >= 2009)
```


```{r}
ggplot(hydro_sub,
       aes(x = factor(month),
           y = hydro_load)) +
    geom_jitter(aes(col = factor(month)),
                size = 0.7,
                alpha = 0.6) +
    geom_boxplot(alpha = 0) +
    scale_y_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Hydrologic Load by month",
         x = "Month",
         y = "Hydrologic Load") +
    theme(legend.position = "none")
```


## Precip by month  

```{r}
ggplot(combnd,
       aes(x = factor(month),
           y = inches)) +
    geom_jitter(aes(col = factor(month)),
                size = 0.7,
                alpha = 0.6) +
    geom_boxplot(alpha = 0) +
    scale_y_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Precip by month",
         x = "Month",
         y = "Monthly Precipitation (inches)") +
    theme(legend.position = "none")
```


## Inside bay vs. not  

```{r}
tbseg2 <- tbseg |> 
    st_transform(crs = st_crs(swfwmdpixels))

fib_nonbay <- st_filter(fib_hydro, st_union(tbseg2), .predicate = st_disjoint)
```

Note, this isn't great spatial coverage. But it does seem more likely that we'd pick up a signal at these stations than if we included all of the ones that are in the bay too.    

```{r}
mapview(fib_nonbay)
```

### Entero by month  

```{r}
ggplot(fib_nonbay,
       aes(x = factor(mo),
           y = val)) +
    geom_jitter(aes(col = factor(mo)),
                size = 0.7,
                alpha = 0.4) +
    geom_boxplot(alpha = 0) +
    scale_y_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero counts by month",
         subtitle = "stations that are not in the bay proper",
         x = "Month",
         y = "Enterococcus") +
    theme(legend.position = "none")
```

Still really not any cycling.... surprising. Maybe a little in Hillsborough Bay.  

### Entero by monthly hydro load  

```{r}
ggplot(fib_nonbay, aes(x = hydro_load,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    # scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. hydrologic load",
         subtitle = "stations that are not in the bay proper",
         x = "Hydrologic Load",
         y = "Enterococcus",
         col = "Month")
```

```{r}
ggplot(fib_nonbay, aes(x = hydro_load,
                      y = val)) +
    geom_jitter(aes(col = bay_segment),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    # facet_wrap(~bay_segment) +
    labs(title = "Entero vs. hydrologic load; log-log scale",
         subtitle = "stations that are not in the bay proper",
         x = "Hydrologic Load",
         y = "Enterococcus",
         col = "Month")
```


### Entero and precip  

```{r}
ggplot(fib_nonbay, aes(x = inches,
                      y = val)) +
    geom_jitter(aes(col = bay_segment),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    # scale_x_log10() +
    # facet_wrap(~bay_segment) +
    labs(title = "Entero vs. precip",
         subtitle = "stations that are not in the bay proper",
         x = "Precipitation (in)",
         y = "Enterococcus",
         col = "Month")
```

```{r}
ggplot(fib_nonbay, aes(x = inches,
                      y = val)) +
    geom_jitter(aes(col = bay_segment),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    # scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. precip",
         subtitle = "stations that are not in the bay proper",
         x = "Precipitation (in)",
         y = "Enterococcus",
         col = "Month")

ggplot(fib_nonbay, aes(x = inches,
                      y = val)) +
    geom_jitter(aes(col = bay_segment),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. precip; log-log scale",
         subtitle = "stations that are not in the bay proper",
         x = "Precipitation (in)",
         y = "Enterococcus",
         col = "Month")
```

Changing the x-axis between log and 'regular' really changes how it looks.  

### Lagged precip  

```{r}
ggplot(fib_nonbay, aes(x = lag1_prcp,
                      y = val)) +
    geom_jitter(aes(col = bay_segment),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    # scale_x_log10() +
    # facet_wrap(~bay_segment) +
    labs(title = "Entero vs. lagged precip",
         subtitle = "stations that are not in the bay proper",
         x = "Previous Month's Precipitation (in)",
         y = "Enterococcus",
         col = "Month")
```

```{r}
ggplot(fib_nonbay, aes(x = lag1_prcp,
                      y = val)) +
    geom_jitter(aes(col = as.factor(mo)),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    # scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. lagged precip",
         subtitle = "stations that are not in the bay proper",
         x = "Previous Month's Precipitation (in)",
         y = "Enterococcus",
         col = "Month")

ggplot(fib_nonbay, aes(x = lag1_prcp,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. lagged precip; log-log scale",
         subtitle = "stations that are not in the bay proper",
         x = "Precipitation (in)",
         y = "Enterococcus",
         col = "Month")
```


## A few individual sites  

OrgID 21FLHILL_WQX
Stations 22, 136, 036, 054



```{r}
fib_tiny <- fib_hydro |> 
    dplyr::filter(OrgID == "21FLHILL_WQX",
           Station %in% c("22", "136", "036", "054"))
mapview(fib_tiny)
```

### Time Series  

```{r}
fib_long <- fib_tiny |> 
    mutate(Date = lubridate::ymd(paste(yr, mo, "01"))) |> 
    rename(Entero = val) |> 
    select(OrgID, Station, Date, Entero, bay_segment, inches, hydro_load, lag1_prcp, lag1_hydro, geometry) |> 
    pivot_longer(c(Entero, inches, hydro_load, lag1_prcp, lag1_hydro),
                 names_to = "variable",
                 values_to = "value")
```

```{r}
ggplot(fib_long, aes(x = Date, y = value, col = Station)) +
    geom_point() +
    geom_line() +
    scale_y_log10() +
    facet_wrap(~variable, scales = "free_y",
               ncol = 1)
```

Well well well, we have at least one disappearing station, and it's one where enterococcus was generally high.  

```{r}
ggplot(fib_tiny, aes(x = inches, y = val, col = Station, group = Station)) +
    geom_point(size = 2, alpha = 0.7) +
    scale_y_log10() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(x = "Precip",
         y = "Entero")
```

And really it looks like values within a station are sort of within their own range, different than other stations, as opposed to really correlating with precip (at least these few stations, on this graph).  Maybe a bit of a relationship if we force it to linear rather than loess.  

```{r}
ggplot(fib_tiny, aes(x = lag1_prcp, y = val, col = Station, group = Station)) +
    geom_point(size = 2, alpha = 0.7) +
    scale_y_log10() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(x = "Lagged Precip",
         y = "Entero")
```

A little different, but not much.  


```{r}
ggplot(fib_tiny, aes(x = hydro_load, y = val, col = Station, group = Station)) +
    geom_point(size = 2, alpha = 0.7) +
    scale_y_log10() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(x = "Hydro load",
         y = "Entero")
```


## Daily-level precipitation  

Incorporate day-of-sampling, day + day before (24 hrs), day + 2 days before (48 hrs), and day + week before (7-day)  

Working with `fib_daily_prcp` data frame.  

```{r}
fib_daily_nonbay <- st_filter(fib_daily_prcp, st_union(tbseg2), .predicate = st_disjoint)

fib_daily_long <- fib_daily_prcp |> 
    pivot_longer(inches:daily_7d,
                 names_to = "time_period",
                 values_to = "total") |> 
    mutate(time_period = case_match(time_period,
                                    "inches" ~ "day of sample",
                                    "daily_24hrs" ~ "sample + previous day",
                                    "daily_48hrs" ~ "sample + 2 days",
                                    "daily_7d" ~ "previous week"))
fib_daily_long_nonbay <- fib_daily_long |> 
    filter(Station %in% fib_daily_nonbay$Station)
```

```{r}
ggplot(fib_daily_long, aes(x = total, y = val)) +
    geom_point(aes(col = time_period),
               alpha = 0.4) +
    facet_wrap(~BAY_SEG) +
    scale_x_log10() +
    scale_y_log10() +
    scale_color_okabeito() +
    labs(title = "All stations",
         subtitle = "note log-log scale",
         x = "Precip (inches)",
         y = "Enterococcus",
         col = "time period of precip total")
```

```{r}
ggplot(fib_daily_long, aes(x = total, y = val)) +
    geom_point(aes(col = BAY_SEG),
               alpha = 0.4) +
    facet_wrap(~time_period) +
    scale_x_log10() +
    scale_y_log10() +
    scale_color_okabeito() +
    labs(title = "All stations",
         subtitle = "note log-log scale",
         x = "Precip (inches)",
         y = "Enterococcus",
         col = "Bay Segment")
```

```{r}
ggplot(fib_daily_long_nonbay, aes(x = total, y = val)) +
    geom_point(aes(col = time_period),
               alpha = 0.4) +
    facet_wrap(~BAY_SEG) +
    scale_x_log10() +
    scale_y_log10() +
    scale_color_okabeito() +
    labs(title = "Non-bay stations",
         subtitle = "note log-log scale",
         x = "Precip (inches)",
         y = "Enterococcus",
         col = "time period of precip total")
```

```{r}
ggplot(fib_daily_long_nonbay, aes(x = total, y = val)) +
    geom_point(aes(col = BAY_SEG),
               alpha = 0.4) +
    facet_wrap(~time_period) +
    scale_x_log10() +
    scale_y_log10() +
    scale_color_okabeito() +
    labs(title = "Non-bay stations",
         subtitle = "note log-log scale",
         x = "Precip (inches)",
         y = "Enterococcus",
         col = "Bay Segment")
```

### Calculate correlations  

Spearman correlations  

#### All stations, all bay segments combined  

```{r}
all_stns <- fib_daily_prcp |> 
    st_drop_geometry() |> 
    select(val, inches:daily_7d) |> 
    set_names(c("Entero", "sample_day_precip",
                "sample_and_24hrs",
                "sample_and_48hrs",
                "previous_week"))
```

```{r}
corr_all <- cor(all_stns,
    use = "pairwise.complete.obs",
    method = "spearman")

knitr::kable(corr_all,
             caption = "Spearman correlation coefficients",
             digits = 3)
knitr::kable(cor.mtest(corr_all)$p,
             caption = "p-values of Spearman correlations",
             digits = 3)

```


#### Non-bay stations, all bay segments combined  

```{r}
nonbay_stns <- fib_daily_nonbay |> 
    st_drop_geometry() |> 
    select(val, inches:daily_7d) |> 
    set_names(c("Entero", "sample_day_precip",
                "sample_and_24hrs",
                "sample_and_48hrs",
                "previous_week"))
```

```{r}
corr_all <- cor(nonbay_stns,
    use = "pairwise.complete.obs",
    method = "spearman")

knitr::kable(corr_all,
             caption = "Spearman correlation coefficients",
             digits = 3)
knitr::kable(cor.mtest(corr_all)$p,
             caption = "p-values of Spearman correlations",
             digits = 3)
```

#### Non-bay stations, by bay segment  

```{r}
options(scipen = 999)

my_cor <- function(df, x){
    cor(df[["val"]], df[[x]],
        method = "spearman",
        use = "pairwise.complete.obs")
}

my_cor.p <- function(df, x){
    cor.test(df[["val"]], df[[x]],
        method = "spearman",
        use = "pairwise.complete.obs")$p.value
}
```

```{r}
nonbay_stns2 <- fib_daily_nonbay |> 
    st_drop_geometry() |> 
    filter(BAY_SEG != "Manatee River") |> # no values, messing things up
    select(BAY_SEG, val, inches:daily_7d) |> 
    summarize(.by = BAY_SEG,
              cor.sampleday = my_cor(.data, "inches"),  # this could all be more elegant
              pval.sampleday = my_cor.p(.data, "inches"),
              cor.24hrs = my_cor(.data, "daily_24hrs"),
              pval.24hrs = my_cor.p(.data, "daily_24hrs"),
              cor.48hrs = my_cor(.data, "daily_48hrs"),
              pval.48hrs = my_cor.p(.data, "daily_48hrs"),
              cor.7d = my_cor(.data, "daily_7d"),
              pval.7d = my_cor.p(.data, "daily_7d"))

knitr::kable(nonbay_stns2,
             caption = "Spearman Correlations and p-values for Entero~rainfall by bay segment",
             digits = 3)    
```

## Wet vs. Dry Season  

Wet season is June-Sept. Dry is the rest.  

```{r}
fib_daily_long_nonbay |> 
    filter(time_period == "sample + 2 days") |> 
ggplot(aes(x = total, y = val, shape = Season)) +
    geom_point(aes(col = Season),
               alpha = 0.5,
               size = 2) +
    facet_wrap(~BAY_SEG) +
    scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
                  labels = c("0.001", "0.01", "0.1", "1")) +
    scale_y_log10() +
    scale_color_muted() +
    labs(title = "Non-bay stations, previous 48 hrs",
         subtitle = "note log-log scale",
         x = "Precip (inches)",
         y = "Enterococcus") +
    theme(axis.text.x = element_text(angle = 40, 
                                     hjust = 1,
                                     vjust = 1))


fib_daily_long_nonbay |> 
    filter(time_period == "sample + 2 days") |> 
ggplot(aes(x = total, y = val, shape = Season)) +
    geom_point(aes(col = Season),
               alpha = 0.5,
               size = 2) +
    facet_wrap(~BAY_SEG) +
    scale_y_log10() +
    scale_color_muted() +
    labs(title = "Non-bay stations, previous 48 hrs",
         subtitle = "note log10 y-axis; x-axis unscaled",
         x = "Precip (inches)",
         y = "Enterococcus") 
```




```{r}
nonbay_stns3 <- fib_daily_nonbay |> 
    st_drop_geometry() |> 
    filter(BAY_SEG != "Manatee River") |> # no values, messign things up
    select(BAY_SEG, Season, val, inches:daily_7d) |> 
    summarize(.by = c(BAY_SEG, Season),
              cor.sampleday = my_cor(.data, "inches"),  # this could all be more elegant
              pval.sampleday = my_cor.p(.data, "inches"),
              cor.24hrs = my_cor(.data, "daily_24hrs"),
              pval.24hrs = my_cor.p(.data, "daily_24hrs"),
              cor.48hrs = my_cor(.data, "daily_48hrs"),
              pval.48hrs = my_cor.p(.data, "daily_48hrs"),
              cor.7d = my_cor(.data, "daily_7d"),
              pval.7d = my_cor.p(.data, "daily_7d"))

knitr::kable(nonbay_stns3,
             caption = "Spearman Correlations and p-values for Entero~rainfall by bay segment",
             digits = 3)    
```


## ECDFs  

Empirical cumulative distribution functions  

### Freshwater  

Differences due to precipitation may show up better in freshwater than estuarine waters; I'll explore that a little bit here just to see what it looks like.    

```{r}
fib_fresh <- fib_dat |> 
    filter(class == "Fresh",
           var == "ecocci") |> 
    mutate(Season = case_when(mo %in% c(6:9) ~ "Wet",
                              .default = "Dry"))

fib_fresh_baysegs <- fib_fresh |> 
    st_as_sf(coords = c("Longitude", "Latitude"),
             crs = "WGS84",
             remove = FALSE) |>    # WGS84 is default for handheld GPS units
    st_transform(crs = st_crs(swfwmdpixels)) |> 
    st_intersection(bay_segs) |> 
    mutate(bay_segment = case_match(BAY_SEG,
                                    "Boca Ciega Bay" ~ "Remainder Lower Tampa Bay",
                                    "Manatee River" ~ "Remainder Lower Tampa Bay",
                                    .default = BAY_SEG))

fib_fresh_hydro <- left_join(fib_fresh_baysegs, combnd,
                       by = c("yr" = "year",
                              "mo" = "month",
                              "bay_segment"))
```

```{r}
ggplot(fib_fresh_hydro, aes(x = val, col = Season)) +
    geom_density() +
    scale_x_log10() +
    facet_wrap(~BAY_SEG, scales = "free_y") +
    labs(title = "Distribution of Enterococcus concentrations at Freshwater stations",
         subtitle = "note log10 x-axis",
         x = "Enterococcus count",
         y = "Density")
```


```{r}
ggplot(fib_fresh_hydro, aes(x = val, col = Season)) +
    stat_ecdf() +
    scale_x_log10() +
    facet_wrap(~BAY_SEG) +
    labs(title = "ECDF for freshwater Enterococcus samples by season and bay segment",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```

```{r}
ggplot(fib_fresh_hydro, aes(x = val, col = Season, group = mo)) +
    stat_ecdf() +
    scale_x_log10() +
    # facet_wrap(~mo) +
    labs(title = "ECDF for freshwater Enterococcus samples by month",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```




### Estuarine

```{r}
fib_nonbay <- fib_nonbay |> 
    mutate(Season = case_when(mo %in% c(6:9) ~ "Wet",
                              .default = "Dry"))
```


```{r}
ggplot(fib_nonbay, aes(x = val, col = Season)) +
    stat_ecdf() +
    scale_x_log10() +
    labs(title = "ECDF for estuarine Enterococcus samples by season across all bay segments",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```

```{r}
ggplot(fib_nonbay, aes(x = val, col = Season)) +
    stat_ecdf() +
    scale_x_log10() +
    facet_wrap(~BAY_SEG) +
    labs(title = "ECDF for estuarine Enterococcus samples by season and bay segment",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```

```{r}
ggplot(fib_nonbay, aes(x = val, col = Season, group = mo)) +
    stat_ecdf() +
    scale_x_log10() +
    # facet_wrap(~mo) +
    labs(title = "ECDF for estuarine Enterococcus samples by month",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```

```{r}
ggplot(fib_nonbay, aes(x = val, col = Season, group = mo)) +
    stat_ecdf() +
    scale_x_log10() +
    facet_wrap(~BAY_SEG) +
    labs(title = "ECDF for estuarine Enterococcus samples by month and bay segment",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```

Not as clear separation in estuarine samples as in freshwater samples, but when all lumped together, there is still some.  

