# Fecal Indicator Bacteria (FIBs) and Hydrology  

This portion contains many exploratory visualizations, as I attempted to find the best way to examine relationships between FIBs - specifically Enterococcus - and hydrological inputs. 

```{r}
library(mapview)
library(sf)
library(tidyverse)
library(tmap)
library(tbeptools)
library(khroma)
library(corrplot)
```

## Data Import  


```{r}
# spatial and precip files
load(here::here("sourced_data", "combined_precip.RData"))
load(here::here("sourced_data", "combined_spatial.RData"))
```


### FIB data import

```{r}
fib_dat <- readRDS(here::here("fib-data", "tbeptools_importwqp.rds"))

fib_sub <- fib_dat |> 
    filter(yr >= 2009,
           class == "Estuary",
           var == "ecocci")

# get those attached to bay segments  
fib_baysegs <- fib_sub |> 
    st_as_sf(coords = c("Longitude", "Latitude"),
             crs = "WGS84",
             remove = FALSE) |>    # WGS84 is default for handheld GPS units
    st_transform(crs = st_crs(swfwmdpixels)) |> 
    st_intersection(bay_segs) |> 
    mutate(bay_segment = case_match(BAY_SEG,
                                    "Boca Ciega Bay" ~ "Remainder Lower Tampa Bay",
                                    "Manatee River" ~ "Remainder Lower Tampa Bay",
                                    .default = BAY_SEG))

fib_hydro <- left_join(fib_baysegs, combnd,
                       by = c("yr" = "year",
                              "mo" = "month",
                              "bay_segment"))

```

### Attach daily precipitation  


```{r}
fib_daily_prcp <- fib_baysegs |> 
    mutate(date = as.Date(SampleTime),
           Season = case_when(mo %in% c(6:9) ~ "Wet",
                              .default = "Dry")) |> 
    left_join(prcp_daily_bayseg,
              by = c("date", "BAY_SEG")) |> 
    filter(!is.na(inches))
```

## All Estuarine Stations - Monthly relationships  

### Entero by monthly hydro load  

```{r}
ggplot(fib_hydro, aes(x = hydro_load,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. hydrologic load; log-log scale",
         x = "Hydrologic Load",
         y = "Enterococcus",
         col = "Month")
```

Interesting..... it's not necessarily the monthly hydrologic loads that matter (though it's clear that the higher hydrologic loads occur in certain months). Could come down to precipitation in some previous time periods.  

Could also be that because the majority of stations are in the bay proper, they're diluted and that's driving the lack-of-relationship. Really need to break this all out by "in-bay" vs. "have some land near the station" (what I later call "non-bay").  

### Lagged hydro load  

```{r}
ggplot(fib_hydro, aes(x = lag1_hydro,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. lagged hydrologic load; log-log scale",
         x = "Hydrologic Load in previous month",
         y = "Enterococcus",
         col = "Month")
```

Nothing new.  

First a little more exploration of entero concentrations by time of year.  

### Entero by monthly precip  

```{r}
ggplot(fib_hydro, aes(x = inches,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.3) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. Monthly Precip",
         x = "Precipitation (in)",
         y = "Enterococcus",
         col = "Month")
```


We see a bit more of a relationship with precip than hydrologic load, at least in the higher values, but nothing clear or strong. What stands out most is that higher precipitation amounts tend to occur in certain months.   


### Previous month's precip  


```{r}
ggplot(fib_hydro, aes(x = lag1_prcp,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.3) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. Lagged Precip",
         x = "Precipitation (in) in previous month",
         y = "Enterococcus",
         col = "Month")
```

Not very different.  


### Entero by month  

```{r}
ggplot(fib_hydro,
       aes(x = factor(mo),
           y = val)) +
    geom_jitter(aes(col = factor(mo)),
                size = 0.7,
                alpha = 0.4) +
    geom_boxplot(alpha = 0) +
    scale_y_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero counts by month",
         x = "Month",
         y = "Enterococcus") +
    theme(legend.position = "none")
```

No clear seasonal patterns in Enterococcus; as a reminder, all stations are included here, and patterns may emerge when we look only at the more land-influenced stations.  

### Hydro load by month  

Only since 2009, to keep data frame size under control.  

```{r}
hydro_sub <- hydro |> 
    dplyr::filter(year >= 2009)
```


```{r}
ggplot(hydro_sub,
       aes(x = factor(month),
           y = hydro_load)) +
    geom_jitter(aes(col = factor(month)),
                size = 0.7,
                alpha = 0.6) +
    geom_boxplot(alpha = 0) +
    scale_y_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Hydrologic Load by month",
         x = "Month",
         y = "Hydrologic Load") +
    theme(legend.position = "none")
```

Clear seasonality.  

### Precip by month  

```{r}
ggplot(combnd,
       aes(x = factor(month),
           y = inches)) +
    geom_jitter(aes(col = factor(month)),
                size = 0.7,
                alpha = 0.6) +
    geom_boxplot(alpha = 0) +
    scale_y_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Precip by month",
         x = "Month",
         y = "Monthly Precipitation (inches)") +
    theme(legend.position = "none")
```

Again, clear seasonality.  


## "Non-bay" estuarine stations - Monthly relationships  

The following stations are those that do not overlap with the `tbeptools::tbseg` layer - and so are not in the bay proper. For lack of a better term, I refer to these as "non-bay" stations. The graphics that follow are generally the same as those above, but on this subset of stations.    

```{r}
tbseg2 <- tbseg |> 
    st_transform(crs = st_crs(swfwmdpixels))

fib_nonbay <- st_filter(fib_hydro, st_union(tbseg2), .predicate = st_disjoint)
```

Note, this isn't great spatial coverage. But it does seem more likely that we'd pick up a signal at these stations than if we included all of the ones that are in the bay too.    

```{r}
mapview(fib_nonbay)
```

### Entero by month  

```{r}
ggplot(fib_nonbay,
       aes(x = factor(mo),
           y = val)) +
    geom_jitter(aes(col = factor(mo)),
                size = 0.7,
                alpha = 0.4) +
    geom_boxplot(alpha = 0) +
    scale_y_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero counts by month",
         subtitle = "stations that are not in the bay proper",
         x = "Month",
         y = "Enterococcus") +
    theme(legend.position = "none")
```

Still really not any cycling.... surprising. There might be a little bit of a signal in Hillsborough Bay.  

### Entero by monthly hydro load  

```{r}
ggplot(fib_nonbay, aes(x = hydro_load,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    # scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. hydrologic load",
         subtitle = "stations that are not in the bay proper",
         x = "Hydrologic Load",
         y = "Enterococcus",
         col = "Month")
```

```{r}
ggplot(fib_nonbay, aes(x = hydro_load,
                      y = val)) +
    geom_jitter(aes(col = bay_segment),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    # facet_wrap(~bay_segment) +
    labs(title = "Entero vs. hydrologic load; log-log scale",
         subtitle = "stations that are not in the bay proper",
         x = "Hydrologic Load",
         y = "Enterococcus",
         col = "Bay Segment")
```


### Entero and precip  

```{r}
ggplot(fib_nonbay, aes(x = inches,
                      y = val)) +
    geom_jitter(aes(col = bay_segment),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    # scale_x_log10() +
    # facet_wrap(~bay_segment) +
    labs(title = "Entero vs. precip",
         subtitle = "stations that are not in the bay proper",
         x = "Precipitation (in)",
         y = "Enterococcus",
         col = "Bay Segment")
```

```{r}
ggplot(fib_nonbay, aes(x = inches,
                      y = val)) +
    geom_jitter(aes(col = bay_segment),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    # scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. precip",
         subtitle = "stations that are not in the bay proper",
         x = "Precipitation (in)",
         y = "Enterococcus",
         col = "Bay Segment") +
    theme(legend.position = "none")

ggplot(fib_nonbay, aes(x = inches,
                      y = val)) +
    geom_jitter(aes(col = bay_segment),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. precip; log-log scale",
         subtitle = "stations that are not in the bay proper",
         x = "Precipitation (in)",
         y = "Enterococcus",
         col = "Bay Segment") +
    theme(legend.position = "none")
```

Changing the x-axis between log and 'regular' really changes how it looks.  

### Lagged precip  

```{r}
ggplot(fib_nonbay, aes(x = lag1_prcp,
                      y = val)) +
    geom_jitter(aes(col = bay_segment),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    # scale_x_log10() +
    # facet_wrap(~bay_segment) +
    labs(title = "Entero vs. lagged precip",
         subtitle = "stations that are not in the bay proper",
         x = "Previous Month's Precipitation (in)",
         y = "Enterococcus",
         col = "Bay Segment")
```

```{r}
ggplot(fib_nonbay, aes(x = lag1_prcp,
                      y = val)) +
    geom_jitter(aes(col = as.factor(mo)),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    # scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. lagged precip",
         subtitle = "stations that are not in the bay proper",
         x = "Previous Month's Precipitation (in)",
         y = "Enterococcus",
         col = "Month")

ggplot(fib_nonbay, aes(x = lag1_prcp,
                      y = val)) +
    geom_jitter(aes(col = factor(mo)),
               size = 2,
               alpha = 0.5) +
    geom_smooth() +
    geom_rug(col = "gray40") +
    scale_y_log10() +
    scale_x_log10() +
    facet_wrap(~bay_segment) +
    labs(title = "Entero vs. lagged precip; log-log scale",
         subtitle = "stations that are not in the bay proper",
         x = "Precipitation (in)",
         y = "Enterococcus",
         col = "Month")
```


## A few individual sites  

I wanted to take a deeper dive into some stations, and chose a few from the map that seemed like they were in interesting locations.  

OrgID 21FLHILL_WQX
Stations 22, 136, 036, 054



```{r}
fib_tiny <- fib_hydro |> 
    dplyr::filter(OrgID == "21FLHILL_WQX",
           Station %in% c("22", "136", "036", "054"))
mapview(fib_tiny)
```

### Time Series  

```{r}
fib_long <- fib_tiny |> 
    mutate(Date = lubridate::ymd(paste(yr, mo, "01"))) |> 
    rename(Entero = val) |> 
    select(OrgID, Station, Date, Entero, bay_segment, inches, hydro_load, lag1_prcp, lag1_hydro, geometry) |> 
    pivot_longer(c(Entero, inches, hydro_load, lag1_prcp, lag1_hydro),
                 names_to = "variable",
                 values_to = "value")
```

```{r}
ggplot(fib_long, aes(x = Date, y = value, col = Station)) +
    geom_point() +
    geom_line() +
    scale_y_log10() +
    facet_wrap(~variable, scales = "free_y",
               ncol = 1)
```

Hm, at least one station stopped being sampled, and it's one where enterococcus was generally high. When doing report cards, will need to make sure we track individual stations or account for variability in sampling in a different way.  

```{r}
ggplot(fib_tiny, aes(x = inches, y = val, col = Station, group = Station)) +
    geom_point(size = 2, alpha = 0.7) +
    scale_y_log10() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(x = "Precip",
         y = "Entero")
```

And really it looks like values within a station are sort of within their own range, different than other stations, as opposed to really correlating with precip (at least these few stations, on this graph).  Maybe a bit of a relationship if we force it to linear rather than loess.  

```{r}
ggplot(fib_tiny, aes(x = lag1_prcp, y = val, col = Station, group = Station)) +
    geom_point(size = 2, alpha = 0.7) +
    scale_y_log10() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(x = "Lagged Precip",
         y = "Entero")
```

A little different, but not much.  


```{r}
ggplot(fib_tiny, aes(x = hydro_load, y = val, col = Station, group = Station)) +
    geom_point(size = 2, alpha = 0.7) +
    scale_y_log10() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(x = "Hydro load",
         y = "Entero")
```


## Daily-level precipitation  

There weren't many relationships with monthly precipitation, but could be more with a finer scale time period. This section incorporates the time periods: day-of-sampling, day + day before (24 hrs), day + 2 days before (48 hrs), and day + week before (7-day)  


```{r}
fib_daily_nonbay <- st_filter(fib_daily_prcp, st_union(tbseg2), .predicate = st_disjoint)

fib_daily_long <- fib_daily_prcp |> 
    pivot_longer(inches:daily_7d,
                 names_to = "time_period",
                 values_to = "total") |> 
    mutate(time_period = case_match(time_period,
                                    "inches" ~ "day of sample",
                                    "daily_24hrs" ~ "sample + previous day",
                                    "daily_48hrs" ~ "sample + 2 days",
                                    "daily_7d" ~ "previous week"))
fib_daily_long_nonbay <- fib_daily_long |> 
    filter(Station %in% fib_daily_nonbay$Station)
```

```{r}
ggplot(fib_daily_long, aes(x = total, y = val)) +
    geom_point(aes(col = time_period),
               alpha = 0.4) +
    facet_wrap(~BAY_SEG) +
    scale_x_log10() +
    scale_y_log10() +
    scale_color_okabeito() +
    labs(title = "All stations",
         subtitle = "note log-log scale",
         x = "Precip (inches)",
         y = "Enterococcus",
         col = "time period of precip total")
```

```{r}
ggplot(fib_daily_long, aes(x = total, y = val)) +
    geom_point(aes(col = BAY_SEG),
               alpha = 0.4) +
    facet_wrap(~time_period) +
    scale_x_log10() +
    scale_y_log10() +
    scale_color_okabeito() +
    labs(title = "All stations",
         subtitle = "note log-log scale",
         x = "Precip (inches)",
         y = "Enterococcus",
         col = "Bay Segment")
```

```{r}
ggplot(fib_daily_long_nonbay, aes(x = total, y = val)) +
    geom_point(aes(col = time_period),
               alpha = 0.4) +
    facet_wrap(~BAY_SEG) +
    scale_x_log10() +
    scale_y_log10() +
    scale_color_okabeito() +
    labs(title = "Non-bay stations",
         subtitle = "note log-log scale",
         x = "Precip (inches)",
         y = "Enterococcus",
         col = "time period of precip total")
```

```{r}
ggplot(fib_daily_long_nonbay, aes(x = total, y = val)) +
    geom_point(aes(col = BAY_SEG),
               alpha = 0.4) +
    facet_wrap(~time_period) +
    scale_x_log10() +
    scale_y_log10() +
    scale_color_okabeito() +
    labs(title = "Non-bay stations",
         subtitle = "note log-log scale",
         x = "Precip (inches)",
         y = "Enterococcus",
         col = "Bay Segment")
```

### Correlations  

Spearman correlations  

#### All stations, all bay segments combined  

```{r}
all_stns <- fib_daily_prcp |> 
    st_drop_geometry() |> 
    select(val, inches:daily_7d) |> 
    set_names(c("Entero", "sample_day_precip",
                "sample_and_24hrs",
                "sample_and_48hrs",
                "previous_week"))
```

```{r}
corr_all <- cor(all_stns,
    use = "pairwise.complete.obs",
    method = "spearman")

knitr::kable(corr_all,
             caption = "Spearman correlation coefficients",
             digits = 3)
knitr::kable(cor.mtest(corr_all)$p,
             caption = "p-values of Spearman correlations",
             digits = 3)

```


#### Non-bay stations, all bay segments combined  

```{r}
nonbay_stns <- fib_daily_nonbay |> 
    st_drop_geometry() |> 
    select(val, inches:daily_7d) |> 
    set_names(c("Entero", "sample_day_precip",
                "sample_and_24hrs",
                "sample_and_48hrs",
                "previous_week"))
```

```{r}
corr_all <- cor(nonbay_stns,
    use = "pairwise.complete.obs",
    method = "spearman")

knitr::kable(corr_all,
             caption = "Spearman correlation coefficients",
             digits = 3)
knitr::kable(cor.mtest(corr_all)$p,
             caption = "p-values of Spearman correlations",
             digits = 3)
```

#### Non-bay stations, by bay segment  

```{r}
options(scipen = 999)

my_cor <- function(df, x){
    cor(df[["val"]], df[[x]],
        method = "spearman",
        use = "pairwise.complete.obs")
}

my_cor.p <- function(df, x){
    cor.test(df[["val"]], df[[x]],
        method = "spearman",
        use = "pairwise.complete.obs")$p.value
}
```

```{r}
nonbay_stns2 <- fib_daily_nonbay |> 
    st_drop_geometry() |> 
    filter(BAY_SEG != "Manatee River") |> # no values, messing things up
    select(BAY_SEG, val, inches:daily_7d) |> 
    summarize(.by = BAY_SEG,
              cor.sampleday = my_cor(.data, "inches"),  # this could all be more elegant
              pval.sampleday = my_cor.p(.data, "inches"),
              cor.24hrs = my_cor(.data, "daily_24hrs"),
              pval.24hrs = my_cor.p(.data, "daily_24hrs"),
              cor.48hrs = my_cor(.data, "daily_48hrs"),
              pval.48hrs = my_cor.p(.data, "daily_48hrs"),
              cor.7d = my_cor(.data, "daily_7d"),
              pval.7d = my_cor.p(.data, "daily_7d"))

knitr::kable(nonbay_stns2,
             caption = "Spearman Correlations and p-values for Entero~rainfall by bay segment",
             digits = 3)    
```

## Wet vs. Dry Season  

Wet season is June-Sept. Dry is the rest.  

```{r}
fib_daily_long_nonbay |> 
    filter(time_period == "sample + 2 days") |> 
ggplot(aes(x = total, y = val, shape = Season)) +
    geom_point(aes(col = Season),
               alpha = 0.5,
               size = 2) +
    facet_wrap(~BAY_SEG) +
    scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
                  labels = c("0.001", "0.01", "0.1", "1")) +
    scale_y_log10() +
    scale_color_muted() +
    labs(title = "Non-bay stations, previous 48 hrs",
         subtitle = "note log-log scale",
         x = "Precip (inches)",
         y = "Enterococcus") +
    theme(axis.text.x = element_text(angle = 40, 
                                     hjust = 1,
                                     vjust = 1))


fib_daily_long_nonbay |> 
    filter(time_period == "sample + 2 days") |> 
ggplot(aes(x = total, y = val, shape = Season)) +
    geom_point(aes(col = Season),
               alpha = 0.5,
               size = 2) +
    facet_wrap(~BAY_SEG) +
    scale_y_log10() +
    scale_color_muted() +
    labs(title = "Non-bay stations, previous 48 hrs",
         subtitle = "note log10 y-axis; x-axis unscaled",
         x = "Precip (inches)",
         y = "Enterococcus") 
```




```{r}
nonbay_stns3 <- fib_daily_nonbay |> 
    st_drop_geometry() |> 
    filter(BAY_SEG != "Manatee River") |> # no values, messign things up
    select(BAY_SEG, Season, val, inches:daily_7d) |> 
    summarize(.by = c(BAY_SEG, Season),
              cor.sampleday = my_cor(.data, "inches"),  # this could all be more elegant
              pval.sampleday = my_cor.p(.data, "inches"),
              cor.24hrs = my_cor(.data, "daily_24hrs"),
              pval.24hrs = my_cor.p(.data, "daily_24hrs"),
              cor.48hrs = my_cor(.data, "daily_48hrs"),
              pval.48hrs = my_cor.p(.data, "daily_48hrs"),
              cor.7d = my_cor(.data, "daily_7d"),
              pval.7d = my_cor.p(.data, "daily_7d"))

knitr::kable(nonbay_stns3,
             caption = "Spearman Correlations and p-values for Entero~rainfall by bay segment",
             digits = 3)    
```


## ECDFs  

Empirical cumulative distribution functions - to explore the general distribution of Enterococcus in wet vs. dry season, in a different way.   

### Freshwater  

Differences due to precipitation may show up better in freshwater than estuarine waters; I'll explore that a little bit here just to see what it looks like.    

```{r}
fib_fresh <- fib_dat |> 
    filter(class == "Fresh",
           var == "ecocci") |> 
    mutate(Season = case_when(mo %in% c(6:9) ~ "Wet",
                              .default = "Dry"))

fib_fresh_baysegs <- fib_fresh |> 
    st_as_sf(coords = c("Longitude", "Latitude"),
             crs = "WGS84",
             remove = FALSE) |>    # WGS84 is default for handheld GPS units
    st_transform(crs = st_crs(swfwmdpixels)) |> 
    st_intersection(bay_segs) |> 
    mutate(bay_segment = case_match(BAY_SEG,
                                    "Boca Ciega Bay" ~ "Remainder Lower Tampa Bay",
                                    "Manatee River" ~ "Remainder Lower Tampa Bay",
                                    .default = BAY_SEG))

fib_fresh_hydro <- left_join(fib_fresh_baysegs, combnd,
                       by = c("yr" = "year",
                              "mo" = "month",
                              "bay_segment"))
```

```{r}
ggplot(fib_fresh_hydro, aes(x = val, col = Season)) +
    geom_density() +
    scale_x_log10() +
    facet_wrap(~BAY_SEG, scales = "free_y") +
    labs(title = "Distribution of Enterococcus concentrations at Freshwater stations",
         subtitle = "note log10 x-axis",
         x = "Enterococcus count",
         y = "Density")
```


```{r}
ggplot(fib_fresh_hydro, aes(x = val, col = Season)) +
    stat_ecdf() +
    scale_x_log10() +
    facet_wrap(~BAY_SEG) +
    labs(title = "ECDF for freshwater Enterococcus samples by season and bay segment",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```

```{r}
ggplot(fib_fresh_hydro, aes(x = val, col = Season, group = mo)) +
    stat_ecdf() +
    scale_x_log10() +
    # facet_wrap(~mo) +
    labs(title = "ECDF for freshwater Enterococcus samples by month",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```




### Estuarine

```{r}
fib_nonbay <- fib_nonbay |> 
    mutate(Season = case_when(mo %in% c(6:9) ~ "Wet",
                              .default = "Dry"))
```


```{r}
ggplot(fib_nonbay, aes(x = val, col = Season)) +
    stat_ecdf() +
    scale_x_log10() +
    labs(title = "ECDF for estuarine Enterococcus samples by season across all bay segments",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```

```{r}
ggplot(fib_nonbay, aes(x = val, col = Season)) +
    stat_ecdf() +
    scale_x_log10() +
    facet_wrap(~BAY_SEG) +
    labs(title = "ECDF for estuarine Enterococcus samples by season and bay segment",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```

```{r}
ggplot(fib_nonbay, aes(x = val, col = Season, group = mo)) +
    stat_ecdf() +
    scale_x_log10() +
    # facet_wrap(~mo) +
    labs(title = "ECDF for estuarine Enterococcus samples by month",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```

```{r}
ggplot(fib_nonbay, aes(x = val, col = Season, group = mo)) +
    stat_ecdf() +
    scale_x_log10() +
    facet_wrap(~BAY_SEG) +
    labs(title = "ECDF for estuarine Enterococcus samples by month and bay segment",
         subtitle = "note log10 x-axis",
         x = "Enterococcus value",
         y = "cumulative proportion")
```

Not as clear separation in estuarine samples as in freshwater samples, but when all lumped together, there is still some. Generally the Enterococcus levels are lower in dry than in wet seasons.   

